A flexible integer linear programming formulation for scheduling clinician on-call service in hospitals

### 2019-09-04 paper/paper.discussion.tex
Add CPU/RAM specs + fix up discussion

In this paper, we present a simple, yet flexible, integer linear programming formulation to generate schedules for clinical departments at hospitals.
The challenge in applying ILP to the task of scheduling clinicians lies in the computational complexity of finding an optimal solution. As the size of our problem grows, due to a larger roster of clinicians or more complicated constraints, the time it takes to generate an optimal schedule may grow exponentially.
%The challenge in applying such an approach to this task lies in the fact that ILP is an NP-hard problem. %specify what 'such an approach' and 'this task' are... the first part of the sentence was a bit hard to follow. Define what you mean by NP-hard problem and cite --> i.e. NP-hard = the following sentence about how time to optimal solution grows exponentially? if yes, then instead of 'As such' - say 'That is,...or "An NP-hard problem means that..."
As a result, the vast majority of approaches taken to create schedules for similar scenarios tend to use heuristics in order to find an approximately optimal solution in a shorter time. \\  %cite/reference statement (i.e. the 'vast majority'...) I can't remember if the introduction details the heuristics --> but if yes, then great. if no, then include that review of the prior literature here in discussion. I think as you said - the journals we are submitting to seem to have more of those details in the introduction.

@@

In our simulations, we also found that increasing the number of requests per clinician does not affect the runtime of the algorithm, highlighting the flexibility of the tool in incorporating clinician preferences. Further, we saw that the algorithm can accommodate increasing the time-horizon up to four years with little impact on runtime, which can be applied to departments that generate schedules far in advance. One key limitation we identified was the sensitivity of the runtime to larger numbers of services offered in a single department. Such cases are unlikely to be encountered in the real-world because most clinical departments tend to provide a single service or at most two services by the same roster of clinicians [ref \ref{???}]. One solution to mitigate the runtime issues generated by a larger number of services would be relaxing certain constraints, followed by manual readjustment from the generated schedule. %what do we mean by certain constraints? specify the other constraints?
Thus, next steps include expanding the generalizability of the tool beyond smaller clinical departments to larger departments within and outside of health-care - especially those that provide multiple services in parallel for patients and other clients. Overall, our sensitivity analyses using simulated data provide reassurance that the ILP formulation can be applied to schedule clinicians across real-world variability between clinical departments. %SM - will look for reference and we can ask Kevin Gough too.

>>>>

In this paper, we present a simple, yet flexible, integer linear programming formulation to generate schedules for clinical departments at hospitals.
The challenge in applying ILP to the task of scheduling clinicians lies in the computational complexity of finding an optimal solution. As the size of our problem grows, due to a larger roster of clinicians or more complicated constraints, the time it takes to generate an optimal schedule may grow exponentially.
%The challenge in applying such an approach to this task lies in the fact that ILP is an NP-hard problem. %specify what 'such an approach' and 'this task' are... the first part of the sentence was a bit hard to follow. Define what you mean by NP-hard problem and cite --> i.e. NP-hard = the following sentence about how time to optimal solution grows exponentially? if yes, then instead of 'As such' - say 'That is,...or "An NP-hard problem means that..."
As a result, the majority of approaches taken to create schedules for similar scenarios tend to use heuristics in order to find an approximately optimal solution in a shorter time \cite{burke_state_2004}. \\  %cite/reference statement (i.e. the 'vast majority'...) I can't remember if the introduction details the heuristics --> but if yes, then great. if no, then include that review of the prior literature here in discussion. I think as you said - the journals we are submitting to seem to have more of those details in the introduction.

@@

In our simulations, we also found that increasing the number of requests per clinician does not affect the runtime of the algorithm, highlighting the flexibility of the tool in incorporating clinician preferences. Further, we saw that the algorithm can accommodate increasing the time-horizon up to four years with little impact on runtime, which can be applied to departments that generate schedules far in advance. One key limitation we identified was the sensitivity of the runtime to larger numbers of services offered in a single department. Such cases are unlikely to be encountered in the real-world because most clinical departments tend to provide a single service or at most two services by the same roster of clinicians [ref \ref{???}]. One solution to mitigate the runtime issues generated by a larger number of services would be relaxing the constraints preventing assignment of consecutive blocks, followed by manual readjustment from the generated schedule. %what do we mean by certain constraints? specify the other constraints?
Thus, next steps include expanding the generalizability of the tool beyond smaller clinical departments to larger departments within and outside of health-care - especially those that provide multiple services in parallel for patients and other clients. Overall, our sensitivity analyses using simulated data provide reassurance that the ILP formulation can be applied to schedule clinicians across real-world variability between clinical departments. %SM - will look for reference and we can ask Kevin Gough too.
